Response time of a plasmonic distributed feedback laser in large-signal modulation regime

### 2019-04-01 Response_DFB_Pulse_PRApplied.md

----------------------------------------------------------------------

# Report of Referee A -- NB10105/Nefedkin

----------------------------------------------------------------------

*This paper theoretically studied the ultrafast response time of a plasmonic distributed feedback laser in the large-signal modulation regime. The authors demonstrated the dependence of response limit with*
*pump spot size using both numerical and analytical approaches. The identification on different ultrafast stages and insights on pump-tunable lasing dynamics could facilitate the applications of plasmon lasers in optoelectronic devices.*

**Review**

*However, the manuscript focused on plasmonic nanohole arrays with a specific parameter design (100-nm hole, 470-nm spacing and 100-nm film thickness). Are the conclusions on the critical pump spot size and*
*ultrafast response limit (1 ps) general enough to plasmon nanohole array lasers in different design parameters (film thickness, hole size, spacing)?*

**Response**

**1.** The time response of the plasmonic laser is determined by the Purcell factor and the EM field relaxation rate in plasmonic structure. These characteristics depend on film thickness, hole size, spacing. For
this reason, the values of the critical pump spot size and ultrafast response limit are different for divergent realizations of plasmon nanohole array lasers.

However, our analytical model shows that the existence of the critical pump spot size is related to finite size of an effective laser mode. When the size of pump beam is greater than the size of the effective
laser mode the response time almost does not depend on the size of pump beam. When the size of the pump beam becomes smaller than the size of the effective laser mode, the response time increases with the decrease of the pump beam size. For this reason, the critical pump spot size takes place in different realizations of the plasmonic distributed-feedback lasers.

**Review**

*Also, plasmon lasers based on nanohole arrays and nanoparticle arrays show different lasing mechanisms from surface plasmon polariton and surface lattice resonances, respectively. It is not a fair comparison*
*between the calculated response time of plasmon nanohole arrays and experimental results of nanoparticle arrays to verify the accuracy of theoretical modeling (final paragraph, the introduction section).*

**Response**

**2.** It is true that in our numerical simulations we consider plasmon lasers based on nanohole arrays, though in the work [Torma] the experiment has been conducted for plasmon lasers based on nanoparticle arrays in the different geometry. In this regard, the conformity of the response times obtained in the experiment and in our simulations is not a fair verification of the accuracy of our theoretical modeling. We agree with the referee in this point and have reformulated the last paragraph in the Introduction section by excluding this direct comparison.

**Review**

*In addition, can the authors more explanations and rationales on how experimental conditions such as pump size and pump power density (excitation fluence) are incorporated in the modeling?*

**Response**

**3.** In our modeling we take into account a large amount of atoms, which have different coordinates $\textbf{r}_m$. The coupling constant between $j$th mode of the plasmonic structure and $m$th atom depends on atom's coordinates as
$$
\Omega_{jm} = \Omega_0 e^{i \textbf{k}_j \textbf{r}_m}
\tag{1}
$$

The size of the pump beam and the power density of pumping are incorporated in the model as the initial conditions for population inversions of atoms. The initial value of the population inversion for the atoms located inside the pumping spot is $D_0$. Other atoms have initial population inversions equal to $-1$, i.e. they are in the ground state. The value of $D_0$ is obtained from the pumping power density as
$$
\frac{D_0 + 1}{2} = \frac{P_0 \Delta t}{n_c S l \hbar \omega}
\tag{2}
$$
Here $P_0$ is power of the laser pulse pump; $\Delta t$ is duration of the pump pulse; $n_c$ is the concentration of atoms of the active medium; $S$ is the area of the pump beam; $l$ is the width of the active medium layer; $\hbar \omega$ is the transition frequency of the atom; and $(D_0 + 1)/2$ is the population of the excited state arising due to the pump pulse.

Thus, we assume that the entire pulse energy comes to the excitation of atoms of the active medium. In real experiment, there are additional channels of energy losses, which lead to the decrease of the population inversion of the excited state. These losses can be taken into account as an addition factor in (2). For the simplicity, we do not consider the additional losses and use the estimation (2).

In our simulations the power of the pump pulse is $P_0 \approx $. This value is typical for experiments carried out in with the same plasmonic structure in the same geometry, see [van exter, Beijnum].

Note that the dependence of the response time on the pump beam size is observable in a wide range of $D_0$ values, and a fine tune of the pumping parameters is not necessary to observe the predicted dependency.

**Review**

*Did the authors consider any diffractive coupling between nanoholes in the equations of system dynamics, given that pump size may influence the resonance mode quality (Purcell factor) and EM fields near the holes?*

**Response**

**4.** We take into account diffractive coupling between nanoholes in laser equations. For this purpose, firstly, we find the EM field distribution of the Bloch modes of the plasmonic structure within the coupled-mode theory. Using this framework we allow for the scattering of the surface plasmon-polariton waves with different Bloch wavenumbers on the nanoholes. We use the scattering rates found in Ref. \[V. T. Tenner, A. N. van Delft, M. J. A. de Dood, and M. P. van Exter, *Journal of Optics* 16, 114019 (2014)\] for the plasmonic laser based on Au film perforated by the nanohole arrays. The coupled-mode theory with such scattering rates demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser \[V. T. Tenner, A. N. van Delft, M. J. A. de Dood, and M. P. van Exter, Journal of Optics 16, 114019 (2014)\].

Secondly, we use the obtained EM field distributions to calculate the Rabi constants between the Bloch modes and atoms of the active medium. The eigenfrequencies and the relaxation rates of the Bloch modes enter in the laser equations (1)-(3).

Thus, the parameters and dispersive dependences of the Bloch modes of the plasmonic structure taken into account in the laser equations via the Rabi constants, the eiegnfrequencies, and the relaxation rates.

We added the description of the band-structure of the plasmonic nanoholes array laser in the Section 2 of main part of the manuscript and **<u>Appendixes A and B.</u>**

A change in the pump beam size results in a change in the volume of the collective mode. As a result, the Purcell factor and the rates of the spontaneous and stimulated emissions also change. The strongest change occurs when the diameter of the pump beam is less than the critical value, at which the strongest dependence of the response time on the diameter appears.

Note that in our modeling the diameter of the pump beam is always much larger than the period of the nanohole array. For this reason, we do not take into account the inhomogeneous pumping within the cell of plasmonic structure.

**Review**

*Also, how is the Rabi constant in the modeling related to the Purcell factor of a specific plasmon cavity? More justifications are needed.*

**5.** We added the derivation of the Rabi constant through the parameters of the plasmonic structure and the active medium in the revised version of the manuscript , see Section 2 in main part of the text and the Appendix A.

**Review**

*We have other comments below*

**1.** *On page 4, the authors claimed that “The time independence of the EM field distribution allows us to consider the plasmonic laser as a single mode laser,” but they need to justify how the time evolution of EM field is correlated to single-mode approximation. Is there any effect from high-order diffractive modes for the hole arrays? Why does the EM field distribution slightly change from Fig 3d to 3f?*

**Response**

In our simulations we take into account several hundred of the Bloch modes of the plasmonic structure which form together the EM field distribution in the active medium. However, our simulations show that the resulting the EM field distribution slightly changes in the first and second stage of the output pulse evolution. In other words, there is a dynamical equilibrium between relaxation and propagation of the EM field on the one hand and the stimulated amplification of the EM field in the active medium on the other hand. We use the fact of the constancy of the EM field distribution to construct the single-mode model. We consider that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multimode problem. This assumption allows to derive the single-mode laser equations. The results obtained in the framework of the single-mode model are in good agreement with the results of the multimode model, see Figs. 2(a), 4 and Fig. 5a.

The change of the EM field distribution from Fig 3d to 3f is associated with a stronger decrease of the population inversion in the third stage of the response. This leads to a violation of  the dynamical equilibrium between relaxation, propagation of the EM field and the stimulated amplification of the EM field in the system. As a result, the EM field distribution begins to change, and the intensity of the EM field begins to decrease.

We added the appropriate explanations in the revised version of the manuscript.

**2.** *Regarding the figures, the order of Figure 5b and 5a can be switched since 5b appeared earlier in the main text. In Figure 5b, what are the units of t_delay? Figure 3 needs additional labeling such as scale bar and timing point of chosen snapshots, and the authors can add more explanations in the figure caption such as the meaning of dashed circles.*

**Response**

We have made the required changes. We changed the order of Figs. 5a and 5b. We indicated the units for $t_\text{delay}$ in Fig. 5b. We added scale bar to the Figure 3; we indicated in the figure caption the time
moments corresponding to the intensity field distribution in Figures 3. We described the meaning of dashed circles in the Figure 3.

We also changed the caption to Figure 3.

In the previous version:

"Field distribution of the collective mode when ..."

In a new version:

"Distribution of the EM field intensity on the surface of plasmonic DFB
laser when ..."

**3.** *The authors point out the similarity of Eq. (15) to the time dependence of a superradiant burst, and how that system resembles the current problem and approach (page 6). The parallelism between these two systems is interesting; however, these few sentences seem out of place relative to the major discussion on ultrafast stages of the laser response. We suggest this information be moved to the introduction or conclusion section, and presented as either as potential motivation or future prospect of study.*

**Response**

We placed all mentions to the superradiance problem to the Conclusion section of the manuscript and noted the potential interest in studying the similarity of the observed phenomenon and superradiance.

----------------------------------------------------------------------

# Report of Referee B -- NB10105/Nefedkin

----------------------------------------------------------------------

*In this manuscript, the Authors theoretically analyze the dynamics of a two-dimensional lasing system, consisting of a periodic metallic sheet supporting plasmonic modes, coated with a lasing material. The dynamics of the lasing phenomenon is studied as a function of the size of the pump laser which generates the population inversion in the lasing material. The Authors find that the response time of the laser, i.e. the delay between the input pump pulse and the output laser pulse, decreases with the size of the pump beam.*

*I find that this manuscript is very well written, with a concise and instructive introductory section. The main result of this work, i.e. the dependence of the response time on the pump beam size, is clearly stated. The Authors also motivate why the study of the response time is important for optoelectronic devices. I recommend this paper for publication, but I kindly ask the Authors to address the two points that I raise below.*

**Review**

*My main criticism of the paper concerns the statement that the calculations reveal an "optimal" beam size, yielding the shortest response time. From Fig.2(b) one sees that the shortest response time corresponds to a very flat minimum in the red curve. I doubt the robustness of this minimum as the system's parameters, or the level of approximation of the theory, are changed. I strongly suggest that the Authors rephrase the text, claiming that d\~15um corresponds to the minimum size to achieve a short response time, and not the optimal size.*

**Response**

**1.** We rephrase the text claiming that $d \sim 15 \mu \text{m}​$ corresponds to the minimum size to achieve the shortest response time, rather than the optimal size.

**Review**

*My second criticism is that the parameters used to calculate the results in Fig.2 are not given in the text. It is written that Fig.1 corresponds to the solution of Eqs.(1)-(3), but what is the size of the device, how many modes are considered, i.e. what are the values of the Rabi couplings, etc.? (The parameters are given only for the single-mode approximation.)*

**Response**

**2.** We have indicated the parameters of the plasmonic DFB laser in the main part of the manuscript. We have added a description of the mode structure of the plasmonic laser and a method for calculating the values of the constants (Rabi constant, the relaxation rates, etc.) from the laser equations (1)--(3).

### 2019-04-03 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Modified by AZ and EA

In our paper, we study the dynamics in the large-signal modulation regime of a plasmonic distributed feedback (DFB) laser that consists of a gold film perforated by a periodic array of holes, and is coated with a layer of an active medium.
We examine the behavior of the laser for different sizes of the pump beam.
Combining numerical simulations of the laser dynamics within the framework developed in Ref. \cite{Zyablovsky2017approach,nefedkin2018acsphot} with an analytical evaluation, we show that the response time strongly depends on the size of the pump beam.
This dependence is due to the complex structure of the modes of the DFB laser.
The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump size.
If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
The response time in this case reaches picoseconds and practically does not change with the size of the pump beam.
In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
Consequently, there is a critical minimum diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the shortest response time. This diameter is about the decay length of the EM field in the plasmonic structure and for the plasmonic laser under consideration is $\sim 15 \mu$m. When the size of the pump beam is greater than $d_{\text{beam}}^*$, the response time of the plasmonic laser is $\sim 1$ $\text{ps}$.
The obtained results confirm that plasmonic DFB lasers can have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime. 
Optimization of the size of the pump beam allows an increase in the modulation speed, as well as the energy efficiency of plasmonic DFB laser devices.

@@

{\Omega _{jm}} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar,

@@

where ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom. The electric field `per one quantum' is inversely proportional to the mode volume (see Appendix A). As a result, the Rabi constant increases with the decrease of the mode volume. The Eqs.~(\ref{eq1})--(\ref{eq3}) allow for the dependence of the stimulated and spontaneous emissions rates on the Rabi constant that enables to take into account the Purcell effect of the plasmonic structure. Every atom of the active medium has its coordinates, which determine the Rabi constants between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}). This allows us to take into account the spatial distribution of the pumping and the population inversion within the active medium.

@@

For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes and coated with the layer of the active medium, see Fig.~\ref{fig1}. We consider that the thickness of the Au film is $100$ $\text{nm}$, the diameter of the holes is $160$ $\text{nm}$, and the distance between holes is $470$ $\text{nm}$. The surface of the entire periodic plasmonic structure is $190 \times 190$ $\mu \text{m}^2$. We consider that the transition frequency of the active medium is close to the frequency of the second-order Bragg grating in the plasmonic structure and the linewidth of the active medium is about 100 nm.

@@

In the absence of the holes, the eigenmodes of the plasmonic structure are TM-modes, similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
These modes have an in-plane component of the magnetic field, $H_{\parallel}$, which is perpendicular to the direction of mode's propagation and both out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}. 
The in-plane electric field, $E_{\parallel}$, is in the direction of the mode's propagation. 
The out-of-plane component of the electric field, $E_{\perp}$, is much greater than the in-plane component, $E_{\parallel}$ \cite{TennerJOpt,TennerACSPhot}.
We find the eigenmodes of the plasmonic structure with the holes within the coupled-mode theory with the experimental values of the scattering rates (see~\cite{TennerJOpt} and Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}). Corresponding dispersion curves are shown in \textbf{\underline{in Fig.~.}} The modes corresponding to different dispersion curves have various radiation losses and field distributions inside the unit cell of the plasmonic crystal (see Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}). The far-field radiation from the plasmonic structure arises mainly due to the scattering of $H_{\parallel}$ on the holes~\cite{TennerJOpt}. 
For this reason, the modes that have a minimum value of $H_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $H_{\parallel}$ at the holes.
In the continuous-wave regime, the lasing takes place at the modes with small radiation losses and a high quality factor lying on the 2nd dispersion curve, \textbf{\underline{see Fig.~}} \cite{TennerJOpt,nefedkin2018acsphot}.

In numerical simulations, we take into account several hundreds of the Bloch modes of the plasmonic structure lying on different dispersion curves. The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat. of \cite{nefedkin2018acsphot}.

We consider the response of the system to ultrashort pulse pumping of the active medium.
Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}). We consider that at an initial time the population inversions of the atoms whose coordinates lie inside the pump beam are expressed as
\begin{equation}\label{eq5SM}
{\left(D_m+1\right)/2} = {P_0 \Delta t / \left(n_c S h \hbar \omega \right)}. 
\end{equation}
Here $\left(D_m+1\right)/2$ is a population of the upper level of the atom; $P_0$ is a pump power; $\Delta t$ is a duration of the pump pulse; $n_c$ is a concentration of the active atoms; $S$ is an area of the pump beam; $h$ is the thickness of Au film; $\omega$ is a frequency of the pump pulse. Also, we consider that the population inversions of the atoms whose coordinates lie outside the pump beam are $-1$ at the initial time.

@@

The existence of an optimal size of the pump beam can be explained with the example of a single-mode model. To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
Although in our simulations we take into account several hundreds of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response. In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field and the stimulated amplification of the EM field in the active medium. When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
In the opposite case, when $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
However, in both cases, the spatial distribution of the electric field intensity remains almost unchanged during its temporal evolution.

We use the fact of the constancy of the EM field distribution to construct a single-mode model. To do this, we assume that there is a single collective mode, the EM field distribution of which coincides with the EM field distribution found in the numerical simulations of the multi-mode problem. Furthermore, we assume that the pumped active medium interacts only with such a single collective mode. Within these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
\begin{gather} 
\dot n =  - 2{\gamma _a}n + 2{N_{\text{at}}}{\Omega _R}\varphi \label{ZE1}
\\
\dot \varphi = - \left( {{\gamma _a} + {\gamma _\sigma }} \right)\varphi + {\Omega _R}\left( {2nD + D + 1} \right)/2 \label{ZE2}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4{\Omega _R}\varphi, \label{ZE3}
\end{gather}
where $n$ is the number of photons in the collective mode, $D$ is the mean population inversion of pumped atoms of the active medium, $\varphi$ describes the mean energy flow from the active medium to the collective mode, and $N_{\text{at}}$ is the number of atoms of the active medium lying in the collective mode.
We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.
Thus, the phase relaxation rate is much larger than other relaxation rates, $\gamma_\sigma \gg \gamma_a, \gamma_D$.
For this reason, one can adiabatically eliminate $\varphi$ from Eqs.~(\ref{ZE1})--(\ref{ZE3}) and obtain the conventional rate equations
\begin{gather} 
\dot n =  - {\gamma _a}n + 2{N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
\end{gather}
As initial condition we take $D(0) = 1$ and $n(0) = 0$, i.e., all atoms are inverted and there are no photons in the mode.
The temporal dynamics of such a system is well-known \cite{SiegmanLasers}; it is shown in Fig.~\ref{fig4}.

@@

It has been shown that the laser's response time can be shortened by optimizing the size of the pump beam, as a result of the consequent reduction of the delay time.
The dependence of the response time on the size of the pump beam is for following reason.
The interaction of the laser modes with the pumped active medium results in the emergence of a single collective mode.
If the size of the pump beam exceeds the propagation length of the EM waves, this collective mode is located in the pumped area.
The response time then grows logarithmically with increasing pump beam diameter.
In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.

Thus, there is a critical minimum diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the shortest response time. This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m. The existence of the critical pump beam size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
In turn, if the area of the pump beam is small, there is the opportunity to reduce the size of the device.
Thus, the obtained results pave the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.

It is interesting to note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}). In the plasmonic cavity, the inequality $\gamma_a \gg \gamma_D$ is always hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds. We believe that further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.

@@

>>>>

In our paper, we study the dynamics in the large-signal modulation regime of a plasmonic distributed feedback (DFB) laser that consists of a gold film perforated by a periodic array of holes, and is coated with a layer of an active medium.
We examine the behavior of the laser for different sizes of the pump beam.
Combining numerical simulations of the laser dynamics within the framework developed in Ref. \cite{Zyablovsky2017approach,nefedkin2018acsphot} with an analytical evaluation, we show that the response time strongly depends on the size of the pump beam.
This dependence is due to the complex structure of the modes of the DFB laser.
The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump size.
If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
The response time in this case reaches picoseconds and practically does not change with the size of the pump beam.
In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
Consequently, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time. This diameter is about the decay length of the EM field in the plasmonic structure and for the plasmonic laser under consideration is $\sim 15 \mu$m. When the size of the pump beam is greater than $d_{\text{beam}}^*$, the response time of the plasmonic laser is $\sim 1$ $\text{ps}$.
The obtained results confirm that plasmonic DFB lasers can have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime. 
Optimization of the size of the pump beam allows an increase in the modulation speed, as well as the energy efficiency of plasmonic DFB laser devices.

@@

{\Omega _{jm}} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar.

@@

Here ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom. Every atom of the active medium has its coordinates, which determine the Rabi constants of interaction between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}). This allows us to take into account the spatial distribution of the pumping and the population inversion within the active medium. Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition, $D_m(0)$ for atoms lying inside the pump beam. The value of $D_m(0)$ is proportional to the pumping energy per unit area. The population inversions of the atoms whose coordinates lie outside the pump beam are $-1$ at the initial time. Note that the square of the Rabi constant is inversely proportional of the Bloch mode volume and is proportional of the Purcell factor of the mode (see Appendix).

@@

For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes and coated with the layer of the active medium, see Fig.~\ref{fig1}. We consider that the thickness of the Au film is $100$ $\text{nm}$, the diameter of the holes is $160$ $\text{nm}$, and the distance between holes is $470$ $\text{nm}$. The surface of the entire periodic plasmonic structure is $190 \times 190$ $\mu \text{m}^2$. We consider that the transition frequency of the active medium, $\omega_\sigma$, is close to the frequency of the second Bloch condition in the plasmonic structure, i.e., $k_{pl}(\omega_{\sigma}) \approx G$, where $k_{pl}$ is a wavevector of the mode in the plasmonic structure and $G$ is a reciprocal lattice vector. \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} and the linewidth of the active medium is about 100 nm.

@@

The EM field distribution of the Bloch modes of the plasmonic structure have been found within the coupled-mode theory \cite{TennerJOpt,TennerACSPhot,nefedkin2018acsphot} (see also Appendix) which describes hybridization of modes of the system without holes through scattering on the holes. Scattering rates for these modes have been taken from \cite{TennerJOpt,TennerACSPhot}. The coupled-mode theory with such scattering rates demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole array laser \cite{TennerJOpt,TennerACSPhot}.

In numerical simulations, we take into account $468$ Bloch modes of the plasmonic structure lying on different dispersion curves. The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat. of \cite{nefedkin2018acsphot}.

@@

The existence of an optimal size of the pump beam can be explained with the example of a single-mode model. To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
Although in our simulations we take into account several hundreds of the Bloch modes of the plasmonic structure, which together form the EM field distribution in the active medium, we can see that the EM field distribution practically does not change in the first and second stages of the system response. In other words, there is a dynamical balance between relaxation processes in the plasmonic structure, the propagation of the EM field outside the pumped area and the stimulated amplification of the EM field inside the pumped area. When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
In the opposite case, when $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
However, in both cases, the spatial distribution of the electric field intensity remains almost unchanged during its temporal evolution.

We use the fact of the constancy of the EM field distribution to construct a single-mode model. To do this, we assume that there is a single collective mode which EM field distribution coincides with the EM field distribution found in the numerical simulations of the multi-mode problem. Furthermore, we assume that the pumped active medium interacts only with such a single collective mode. Within these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
\begin{gather} 
\dot n =  - 2{\gamma _a}n + 2{N_{\text{at}}}{\Omega _R}\varphi \label{ZE1}
\\
\dot \varphi = - \left( {{\gamma _a} + {\gamma _\sigma }} \right)\varphi + {\Omega _R}\left( {2nD + D + 1} \right)/2 \label{ZE2}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4{\Omega _R}\varphi, \label{ZE3}
\end{gather}
where $n$ is the number of photons in the collective mode, $D$ is the mean population inversion of pumped atoms of the active medium, $\varphi$ describes the mean energy flow from the active medium to the collective mode, and $N_{\text{at}}$ is the number of atoms of the active medium lying in the collective mode.
We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.
Thus, the phase relaxation rate is much larger than other relaxation rates, $\gamma_\sigma \gg \gamma_a, \gamma_D$.
For this reason, one can adiabatically eliminate $\varphi$ from Eqs.~(\ref{ZE1})--(\ref{ZE3}) and obtain the conventional rate equations
\begin{gather} 
\dot n =  - {\gamma _a}n + 2{N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
\end{gather}
As initial condition we take $D(0) = 1$ and $n(0) = 0$, i.e., all atoms are inverted and there are no photons in the mode.
The temporal dynamics of such a system is well-known \cite{SiegmanLasers}; it is shown in Fig.~\ref{fig4}.

@@

It has been shown that the laser's response time can be shortened by optimizing the size of the pump beam, as a result of the consequent reduction of the delay time.
The dependence of the response time on the size of the pump beam is for following reason.
The interaction of the laser modes with the pumped active medium results in the emergence of a single collective mode.
If the size of the pump beam exceeds the propagation length of the EM waves, this collective mode is located in the pumped area.
The response time then grows logarithmically with increasing pump beam diameter.
In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.
It is interesting to note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}). In the plasmonic cavity, the inequality $\gamma_a \gg \gamma_D$ is always hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds. We believe that further study of the relationship between these problems may be important in the context of creating ultrafast optoelectronic devices.

Thus, there is the diameter of the pump beam, $d_{\text{beam}}^*$, necessary to achieve the short response time. This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m. The existence of the critical pump beam size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
In turn, if the area of the pump beam is small, there is the opportunity to reduce the size of the device.
Thus, the obtained results pave the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.

@@

The functions ${{\bf{e}}_j}\left(\textbf{r}\right)$ and ${{\bf{h}}_j}\left(\textbf{r}\right)$ can be find within the couple-mode theory, see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat of \cite{nefedkin2018acsphot}. In this approach the eigenmodes of the plasmonic structure with holes are expressed as a linear combinations of the eigenmodes of the plasmonic structure without holes \cite{TennerJOpt,TennerACSPhot}. The expansion coefficients are calculated through the scattering matrix which contains the scattering amplitudes at angles $0^{\circ }$, $\pm 90^{\circ }$ and $180^{\circ }$ and effective refractive index of the plasmonic structure \cite{TennerJOpt,TennerACSPhot}.

The eigenmodes of the plasmonic structure without holes are TM-modes, similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
These modes have an in-plane component of the magnetic field, $h_{\parallel}$, which is perpendicular to the direction of mode's propagation and both out-of-plane, $e_{\perp}$, and in-plane, $e_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}. 
The in-plane electric field, $e_{\parallel}$, is in the direction of the mode's propagation. 
The out-of-plane component of the electric field, $e_{\perp}$, is much greater than the in-plane component, $e_{\parallel}$ \cite{TennerJOpt,TennerACSPhot} and so it plays main role at the interaction of the modes with the active medium.

The eigenmodes of the plasmonic structure with the holes we find within the coupled-mode theory with the experimental values of the scattering amplitudes, see~\cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}. The coupled-mode theory with such scattering amplitudes demonstrates a good agreement with the experimental measurement of the band-structure of the plasmonic nanohole arrays laser \cite{TennerJOpt}.
At close to the transition frequency of the active medium there are four different dispersion curves. The eigenmodes corresponding to different dispersion curves have various radiation losses and EM field distributions inside the unit cell of the plasmonic structure (see \cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}). In particularly, at the edge of the band gap when the Bloch wavevectors, $\left(k_{xj}, k_{yj}\right) = \left(0, 0\right)$  the EM field distributions in the modes from the different dispersion curves are

\textbf{Table A}

$\begin{array}{cccc}
\text{The number of the dispersive curve} & e_{\parallel} & \
h_{\parallel} & \
e_ {\perp } \left( e_z \right) \\
\text{1} & \sin (G x)+\sin (G y) & \sin (G x)+\sin (G y) & \
\cos (G x)+\cos (G y) \\
\text{2} & \sin (G x)-\sin (G y) & \sin (G x)-\sin (G y) & \
\cos (G x)-\cos (G y) \\
\text{3} & \cos (G y) & \cos (G y) & \sin (G y) \\
\text{4} & \cos (G x) & \cos (G x) & \sin (G x) \\
\end{array}$
\\
see~\cite{TennerJOpt,TennerACSPhot} and Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}.
